Dynamic Knapsack Problem

### 2021-03-01 Manuscript/Manuscript.tex
started implementation

\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

@@

\subsection{Questions}
\begin{itemize}
	\item  How to incorporate goal variable $uv$ properly into transition, ADP, etc.?
	\item Does the ADP look correct to you so far?
	\item Is it safe to remove $\beta_{Tp}^{uu}$ and $\beta_{Tmdc}^{ps}$ since they are \textit{supposed} to be zero always
\end{itemize}

@@

\subsection{State Space}
State space is defined by current units available for various PPEs for future periods, amount of units already used for various PPEs, current patient waitlist, expected period demand, and number of patients already scheduled
\[  \vec{S}  = (\vec{ue}, \vec{uu}, \vec{pw}, \vec{pe}, \vec{ps}) \]
\begin{itemize}
    \item $\vec{ue} = ue_{tp}$ - Expected unist for period $t$, and PPE $p$
    \item $\vec{uu} = uu_{tp}$ - Used units for period $t$, and PPE $p$
    \item $\vec{pw} = pw_{mdc}$ - Number of patients of complexity $d$, CPU $c$, on a wait list for $m$ periods
    \item $\vec{pe} = pe_{dc}$ - Number of patients of complexity $d$, CPU $c$ expected to arrive this period
    \item $\vec{ps} = ps_{tmdc}$ - Number of patients of complexity $d$, CPU $c$, scheduled to period $t$, who have been on the waitlist for $m$ periods ($m$ of 0 stands for people who have just arrived)
\end{itemize}

@@

\subsubsection{Action Constraints}
\begin{itemize}
    \item Total number of PPE units cannot be exceeded  
    \begin{itemize}
    		\item ($U_{pdc}$ - usage of PPE $p$ per patient difficulty $d$, CPU $c$)
    \end{itemize}
        \begin{equation} 
			\sum_{mdc}(sc_{tmdc})U_{pdc} <= (ue_{tp} - uu_{tp}) + uv_{tp}\ \quad \forall tp
		\end{equation}

@@

	\item number of people scheduled/rescheduled must be consistent
		\begin{equation}  
			\sum_{t'} rsc_{tt'mdc} \le ps_{tmdc} \quad \forall tmdc 
		\end{equation} 
		\begin{equation}  
			\sum_{t} sc_{t0dc} \le pe_{dc} \quad \forall dc 
		\end{equation} 
		\begin{equation}  
			\sum_{t} sc_{tmdc} \le pw_{mdc} \quad \forall m \in \{1...M \}, dc 
		\end{equation}

@@

	\item Amout of expected units of PPE resource for the next time period - $bc_{1dc}$
		\begin{itemize}
			\item let's assume $ued_{md}$ is the random variable that represents the deviation of PPE units from the expectation for the next period only. It follows some uniform distribution.
			\item let's assume $uen_{d}$ is the default value to be used for expected number of PPE units $p$ per period
		\end{itemize}
\end{enumerate}

@@

\textit{\textbf{FROM CONSTRAINT SET \#2 MISSING RESOURCE USAGE DUE TO TRANSITION IN DIFFICULTIES}}
\begin{enumerate}

	\item Transition from $\vec{ue}$ to $\vec{ue'}$ - Expected Units of PPE
		\begin{equation}  
			ue'_{1p} = ue_{2p} + ued_{p}\ \quad \forall p 
		\end{equation} 
		\begin{equation}  
			ue'_{t-1,p} = ue_{tp}\ \quad \forall t \in \{3...T\}, p
		 \end{equation} 
		\begin{equation}  
			ue'_{Tp} = uen_{d} \quad \forall p 
		\end{equation}

@@

@@

	\item $\beta_{tp}^{ue}$
	\begin{equation*}
		E[ued_{p}] = 0 \quad \therefore \quad
		\sum_{p} \beta_{1p}^{ue} (ue_{2p} + ued_{p}) \rightarrow 
		\sum_{p} \beta_{1p}^{ue} (ue_{2p}) \rightarrow
		\sum \beta_{tp}^{ue} uen_{d}
	\end{equation*}
	\begin{equation*}
		\sum_{tp} \beta_{tp}^{ue} ue_{tp} - \gamma (\sum_{tp} \beta_{tp}^{ue} uen_{d}) = 
		\color{red} (1-\gamma) \sum_{tp} \beta_{tp}^{ue} (ue_{tp} - uen_{d})
	\end{equation*}

	\item $\beta_{tp}^{uu}$ - no adjustments for expectation
	\begin{equation*}
		uu_{Tp} = 0 \quad \therefore \quad \beta_{Tp}^{uu} uu_{Tp} = 0
	\end{equation*}
	\begin{equation*}
		\sum_{tp}^{T-1} \beta_{tp}^{uu} uu_{tp} - \gamma \sum_{tp}^{T-1} \beta_{tp}^{uu} 
		\bigg( uu_{t+1p} + \sum_{mdc} (sc_{t+1mdc}U)  - \sum_{tmdc} (rsc_{t+1t'mdc} U) + \sum_{t'mdc} (rsc_{tt'+1mdc} U) \bigg)
	\end{equation*}
	\begin{equation*}
		\color{red} \sum_{tp}^{T-1} \beta_{tp}^{uu} \bigg( 
			uu_{tp}	 - \gamma \Big( uu_{t+1p} + \sum_{mdc} pst_{tmdc} (U_{pd+1c} - U_{pdc}) + 
			U_{pdc} \big( \sum_{mdc} sc_{t+1mdc}  - \sum_{t'mdc} rsc_{t+1t'mdc} + \sum_{tmdc} rsc_{tt'+1mdc} \big) \Big)
		\bigg)
	\end{equation*}

@@

	\item $\beta_{mdc}^{pw}$
	\begin{equation*}
		E[pst_{tmdc}] = ps_{tmdc} pwp_{mdc}		
	\end{equation*}
	{ \color{red} \begin{align*}
		\Bigg( \sum _{t=1 m=1 dc}^{T-1 M-1} \beta_{tmdc}^{ps} & 
			\bigg( ps_{tmdc} - 
				\gamma \Big( ps_{t+1m-1dc}(1-pwp) + ps_{t+1m-1d-1c} pwp - \sum_{t} sc_{t+1m-1dc} + \\ 
				&  - \sum_{t'} rsc_{t+1t'm-1dc} + \sum_{t} rsc_{tt'+1m-1dc} \Big) \bigg) + \\
		\sum _{t=1dc}^{T-1} \beta_{tMdc}^{ps} & 
			\bigg( ps_{tMdc} - 
				\gamma \sum_{M-1}^{M} \Big( ps_{t+1mdc}(1-pwp) + ps_{t+1md-1c} pwp - \sum_{t} sc_{t+1mdc} + \\ 
				&  - \sum_{t'} rsc_{t+1t'mdc} + \sum_{t} rsc_{tt'+1mdc} \Big) \bigg) \Bigg)
	\end{align*} }

@@

\subsection{Dual of ADP LP}
Converting to Dual
\begin{equation}
	\min_{w} \sum_{\vec{S} \vec{A}} w(\vec{S} \vec{A}) c(\vec{S} \vec{A})
\end{equation}
Subject To:
	\begin{equation}
		\sum_{\vec{S}\vec{A}}\vec{w} (1 - \gamma) \ge 1 
	\end{equation}
	\begin{equation}
		\sum_{\vec{S}\vec{A}}\vec{w} \Big( (1-\gamma) (ue_{tp} - uen_{tp}) \Big) \ge E[ue_{tp}] \quad \forall tp
	\end{equation}	
	\begin{multline}
		\sum_{\vec{S}\vec{S}} \vec{w} \bigg( 
			uu_{tp}	 - \gamma \Big( uu_{t+1p} + \sum_{mdc} pst_{tmdc} (U_{pd+1c} - U_{pdc}) + \\ 
			U_{pdc} \big( \sum_{mdc} sc_{t+1mdc}  - \sum_{t'mdc} rsc_{t+1t'mdc} + \sum_{tmdc} rsc_{tt'+1mdc} \big) \Big)
		\bigg) \\
		\quad \forall T \in \{1...T-1\}p
	\end{multline}
	\begin{equation}
		\sum_{\vec{S}\vec{A}} \vec{w} (pe_{dc} - \gamma E[pea_{dc}])  \ge E[pe_{dc}] \quad \forall dc
	\end{equation}
	\begin{equation}
		w \ge 0 \quad \forall w
	\end{equation}

\subsection{Finding most violated constraints of ADP LP}

>>>>

\usepackage{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

@@

\subsection{Questions}
\begin{itemize}
	\item  NEED TO MOVE $uv$ TO STATES AND ADJUST ALL THE RELEVANT EQUATIONS FOR IT
\end{itemize}

@@

\subsection{State Space}
State space is defined by current units available for various PPEs for future periods, amount of units already used for various PPEs, current patient waitlist, expected period demand, and number of patients already scheduled
\[  \vec{S}  = (\vec{ue}, \vec{uu}, \vec{uv}, \vec{pw}, \vec{pe}, \vec{ps}) \]
\begin{itemize}
    \item $\vec{ue} = ue_{tp}$ - Expected unist for period $t$, and PPE $p$
    \item $\vec{uu} = uu_{tp}$ - Used units for period $t$, and PPE $p$
    \item $\vec{uv} = uv_{tp}$ - goal variable, violation on number of resources used for period $t$, of PPE $p$
    \item $\vec{pw} = pw_{mdc}$ - Number of patients of complexity $d$, CPU $c$, on a wait list for $m$ periods
    \item $\vec{pe} = pe_{dc}$ - Number of patients of complexity $d$, CPU $c$ expected to arrive this period
    \item $\vec{ps} = ps_{tmdc}$ - Number of patients of complexity $d$, CPU $c$, scheduled to period $t$, who have been on the waitlist for $m$ periods ($m$ of 0 stands for people who have just arrived)
\end{itemize}

@@

\subsubsection{Action Constraints}
\label{action constraints}
\begin{itemize}
    \item Total number of PPE units cannot be exceeded  
    \begin{itemize}
    		\item ($U_{pdc}$ - usage of PPE $p$ per patient difficulty $d$, CPU $c$)
    \end{itemize}
        \begin{equation} 
			\sum_{mdc}(sc_{tmdc})U_{pdc} <= (ue_{tp} - uu_{tp}) + uv_{tp}\ \quad \forall tp
		\end{equation}

@@

	\item number of people scheduled/rescheduled must be consistent
		\begin{equation}  
			\sum_{t'} rsc_{tt'mdc} \le ps_{tmdc} \quad \forall tmdc 
		\end{equation} 
		\begin{equation}  
			\sum_{t} sc_{t0dc} \le pe_{dc} \quad \forall dc 
		\end{equation} 
		\begin{equation}  
			\sum_{t} sc_{tmdc} \le pw_{mdc} \quad \forall m \in \{1...M \}, dc 
		\end{equation} 
\end{itemize}

@@

	\item Amout of expected units of PPE resource for the next time period - $ue_{1p}$
		\begin{itemize}
			\item let's assume $ued_{d}$ is the random variable that represents the deviation of PPE units from the expectation for the next period only. It follows some uniform distribution.
			\item let's assume $uen_{d}$ is the default value to be used for expected number of PPE units $p$ per period
		\end{itemize}
\end{enumerate}

@@

\begin{enumerate}

	\item Transition from $\vec{ue}$ to $\vec{ue'}$ - Expected Units of PPE
		\begin{equation}  
			ue'_{1p} = ue_{2p} + ued_{p}\ \quad \forall p 
		\end{equation} 
		\begin{equation}  
			ue'_{t-1,p} = ue_{tp}\ \quad \forall t \in \{3...T\}, p
		 \end{equation} 
		\begin{equation}  
			ue'_{Tp} = uen_{p} \quad \forall p 
		\end{equation}

@@

	\item Transition from $\vec{uv}$ to $\vec{uv'}$ - Violation of PPE units
	\begin{equation}
		uv'_{tp} \ge uu'_{tp} - ue'_{tp} \quad \forall tp
	\end{equation}
	\begin{equation}
		uv'_{tp} \ge 0 \quad \forall tp
	\end{equation}
	\begin{equation}
		uv'_{tp} = \max \{0, uu'_{tp} - ue'_{tp} \} \quad \forall tp
	\end{equation}

@@

	\item $\beta_{tp}^{ue}$
	\begin{equation*}
		E[ued_{p}] = 0 \quad \therefore \quad
		\sum_{p} \beta_{1p}^{ue} (ue_{2p} + ued_{p}) \rightarrow 
		\sum_{p} \beta_{1p}^{ue} (ue_{2p}) \rightarrow
		\sum \beta_{tp}^{ue} uen_{p}
	\end{equation*}
	\begin{equation*}
		\sum_{tp} \beta_{tp}^{ue} ue_{tp} - \gamma (\sum_{tp} \beta_{tp}^{ue} uen_{p}) = 
		\color{red} \sum_{tp} \beta_{tp}^{ue} (ue_{tp} - \gamma (uen_{p}))
	\end{equation*}

	\item $\beta_{tp}^{uu}$ - no adjustments for expectation
	\begin{equation*}
		uu_{Tp} = 0 \quad \therefore \quad \beta_{Tp}^{uu} uu_{Tp} = 0
	\end{equation*}
	\begin{align*}
		\sum_{tp}^{T-1} \beta_{tp}^{uu} uu_{tp} - \gamma \sum_{tp}^{T-1} \beta_{tp}^{uu} \bigg( 
					& uu_{t+1p} + \sum_{mdc} (sc_{t+1mdc}U)  - \sum_{tmdc} (rsc_{t+1t'mdc} U) + \sum_{t'mdc} (rsc_{tt'+1mdc} U) + \\ 
					&\sum_{mdc} ps_{tmdc}pwp_{mdc} (U_{pd+1c} - U_{pdc}) \bigg)
	\end{align*}
	{ \color{red} \begin{align*}
		 \Bigg( \sum_{tp}^{T-1} \beta_{tp}^{uu} \bigg( & 
		 	uu_{tp}	 - \gamma \Big( uu_{t+1p} + \sum_{mdc} ps_{tmdc}pwp_{mdc} (U_{pd+1c} - U_{pdc}) + \\
			& U_{pdc} \big( \sum_{mdc} sc_{t+1mdc}  - \sum_{t'mdc} rsc_{t+1t'mdc} + \sum_{tmdc} rsc_{tt'+1mdc} \big) \Big) \bigg) + \\
		\beta_{Tp}^{uu} (uu_{Tp}) \Bigg)
	\end{align*} }

@@

	\item $\beta_{mdc}^{ps}$
	\begin{equation*}
		E[pst_{tmdc}] = ps_{tmdc} pwp_{mdc}		
	\end{equation*}
	{ \color{red} \begin{align*}
		\Bigg( \sum _{t=1 m=1 dc}^{T-1 M-1} \beta_{tmdc}^{ps} & 
			\bigg( ps_{tmdc} - 
				\gamma \Big( ps_{t+1m-1dc}(1-pwp) + ps_{t+1m-1d-1c} pwp - \sum_{t} sc_{t+1m-1dc} \\ 
				&  - \sum_{t'} rsc_{t+1t'm-1dc} + \sum_{t} rsc_{tt'+1m-1dc} \Big) \bigg) + \\
		\sum _{t=1dc}^{T-1} \beta_{tMdc}^{ps} & 
			\bigg( ps_{tMdc} - 
				\gamma \sum_{M-1}^{M} \Big( ps_{t+1mdc}(1-pwp) + ps_{t+1md-1c} pwp - \sum_{t} sc_{t+1mdc} \\ 
				&  - \sum_{t'} rsc_{t+1t'mdc} + \sum_{t} rsc_{tt'+1mdc} \Big) \bigg) + \\ 
		\sum_{mdc} \beta_{Tmdc}^{ps} & 
			\bigg( ps_{Tmdc} \bigg) \Bigg)
	\end{align*} }

@@

\subsection{Dual of ADP LP}
\label{Dual of ADP LP}
Converting to Dual
\begin{equation}
	\min_{w} \sum_{\vec{S} \vec{A}} w(\vec{S} \vec{A}) c(\vec{S} \vec{A})
\end{equation}
Subject To:
\begin{itemize}
	\item $\beta^{0}$ constraints
	\begin{equation}
		\sum_{\vec{S}\vec{A}}\vec{w} (1 - \gamma) = 1 
	\end{equation}

	\item $\beta^{ue}$ constraints	
	\begin{equation}
		\sum_{\vec{S}\vec{A}}\vec{w} \Big(ue_{tp} - \gamma (uen_{tp}) \Big) \ge E[ue_{tp}] \quad \forall tp
	\end{equation}

	\item $\beta^{uu}$ constraints
	\begin{multline}
		\sum_{\vec{S}\vec{A}} \vec{w} \bigg( 
			uu_{tp}	 - \gamma \Big( uu_{t+1p} + \sum_{mdc} ps_{tmdc}pwp (U_{pd+1c} - U_{pdc}) + \\ 
			U_{pdc} \big( \sum_{mdc} sc_{t+1mdc}  - \sum_{t'mdc} rsc_{t+1t'mdc} + \sum_{tmdc} rsc_{tt'+1mdc} \big) \Big)
		\bigg) \ge E[uu_{tp}] \\
		\quad \forall T \in \{1...T-1\}p
	\end{multline}
	\begin{equation}
		\sum_{\vec{S}\vec{A}} \vec{w} (uu_{Tp}) \ge E[uu_{Tp}] \quad \forall p
	\end{equation}

	\item $\beta^{pe}$ constraints
	\begin{equation}
		\sum_{\vec{S}\vec{A}} \vec{w} (pe_{dc} - \gamma E[pea_{dc}])  \ge E[pe_{dc}] \quad \forall dc
	\end{equation}

	\item $\beta^{pw}$ constraints
	\begin{equation}
		  \sum_{\vec{S}\vec{A}}w  ( pw_{1dc} - \gamma (pe_{dc} - \sum_{t} sc_{t0dc})) \ge E[pw_{1dc}] \quad \forall dc
	\end{equation}
	\begin{multline}
		\sum_{\vec{S}\vec{A}}w  \Big( pw_{mdc} - \gamma \big( pw_{m-1dc} (1 - pwp) + pw_{m-1d-1c} pwp - \sum_{t} sc_{tm-1dc} \big) \Big) \\ 
		\ge E[pw_{mdc}]  \quad \forall m \in \{2 ... M-1 \}dc
	\end{multline}
	\begin{equation}
		\sum_{\vec{S}\vec{A}}w  \Big( pw_{Mdc} - \gamma \sum_{M-1}^{M}
			\big( pw_{mdc} (1 - pwp) + pw_{md-1c} pwp - \sum_{t} sc_{tmdc} \big) \Big) \ge E[pw_{Mdc}] \quad \forall dc
	\end{equation}

	\item $\beta^{ps}$ constraints
	\begin{multline}
		\sum_{\vec{S}\vec{A}}w \bigg( ps_{tmdc} - 
				\gamma \Big( ps_{t+1m-1dc}(1-pwp) + ps_{t+1m-1d-1c} pwp - \sum_{t} sc_{t+1m-1dc} \\ 
				 - \sum_{t'} rsc_{t+1t'm-1dc} + \sum_{t} rsc_{tt'+1m-1dc} \Big) \bigg) \\ 
				 \ge E[ps_{tmdc}] \quad \forall t \in \{t...T-1 \} m \in \{1 ... M-1 \} dc
	\end{multline}
	\begin{multline}
		\sum_{\vec{S}\vec{A}}w \bigg( ps_{tMdc} - 
				\gamma \sum_{M-1}^{M} \Big( ps_{t+1mdc}(1-pwp) + ps_{t+1md-1c} pwp - \sum_{t} sc_{t+1mdc} \\ 
				 - \sum_{t'} rsc_{t+1t'mdc} + \sum_{t} rsc_{tt'+1mdc} \Big) \bigg) \\ 
				 \ge E[ps_{tmdc}] \quad \forall t \in \{t...T-1 \} dc
	\end{multline}
	\begin{equation}
		\sum_{\vec{S}\vec{A}} \vec{w} (ps_{Tmdc}) \ge E[ps_{Tmdc}] \quad \forall mdc
	\end{equation}

	\item State-Action constraints
	\begin{equation}
		w \ge 0 \quad \forall w
	\end{equation}
\end{itemize}

\subsection{Finding most violated constraints of ADP LP}
\label{Pricing Problem}
\begin{equation}
	\min_{(\vec{ue}, \vec{uu}, \vec{pw}, \vec{pe}, \vec{ps}) \in S,  (\vec{sc}, \vec{rsc}) \in A } c(\vec{pw},\vec{sc},\vec{rsc},\vec{uv}) - E[V]
\end{equation}
Subject to:
all constraints in section "\nameref{action constraints}"

\subsection{Algorithm for solving}
To get $\vec{\beta}$ values, which will be used to generate an action follow steps below:
\begin{enumerate}
	\item Perform a monte-carlo simulation (following some arbitrary policy) to get $E[ue], E[uu], E[pe], E[pw] E[pw]$ (this will only give an approximation as arbitrary policy will likely be different from ADP policy) (if you wanted to, you could resolve for $\vec{\beta}$ again, but replacing $E[ue], E[uu], E[pe], ...$ with the ones that would be generated through the ADP policy) 
	\item Create an initial feasible set of state-action pairs - $\vec{w}$
	\item Solve model in section "\nameref{Dual of ADP LP}" where each state-action pairs in $\vec{w}$ corresponds to a variable and parameters for all the constraints
	\item Solve model in section "\nameref{Pricing Problem}", where duals from problem in step 3  correspond to $\vec{\beta}$ values. 
		\begin{itemize}
			\item If objective function is less than 0, add solution as a single state-action pair to $\vec{w}$ and go to step 3
			\item If objective function is greater than 0, continue to next step
		\end{itemize}
	\item Duals from problem in step 3 correspond to final $\vec{\beta}$ values
\end{enumerate}

\subsection{Generating an Action}
Once $\vec{\beta}$ values have been approximated, you may use the model below to generate a recommended action for a specific state.
\begin{equation}
	\min_{\vec{A}} c(\vec{S}, \vec{A}) + E[V]
\end{equation}
Subject to:
all constraints in section "\nameref{action constraints}"
